### 2019-10-30 manuscript.Rmd
Replace the text on regression model for mixture model

---
title: "Metabolic Effects of PCSK9 Inhibition With Evolocumab in subjects with elevated Lp(a)"
output: word_document
bibliography: /home/xiang/Documents/Bibliography/lpa.bib
csl: circulation.csl
---

Xiang Zhang, PhD$^1,^2$,
Lotte C.A. Stiekema, MD$^2$,
Erik S.G. Stroes, PhD$^2$,
Albert K. Groen, PhD$^1$

1. Department of Experimental Vascular Medicine, Amsterdam University Medical Center, University of Amsterdam, Meibergdreef 9, 1105 AZ Amsterdam, The Netherlands
2. Human and Animal Physiology, Wageningen University, De Elst 1, 6708 WD Wageningen, The Netherlands
3. Department of Vascular Medicine, Amsterdam University Medical Center, University of Amsterdam, Meibergdreef 9, 1105 AZ Amsterdam, The Netherlands

# Abstract

Background: 
Epidemiological studies show that subjects with elevated lipoprotein(a) [Lp(a)] have a higher risk for developing cardiovascular diseases. 
Inhibition of proprotein convertase subtilisin/kexin type 9 (PCSK9) is one of the limited options to reduce Lp(a). 
In this study, we aimed to investigate metabolic effects of PCSK9 inhibition with evolocumab in subjects with elevated Lp(a).

Methods:
We performed metabolic profiling of plasma samples derived from 30 individuals (placebo N = 14; evolocumab N = 16). 
By applying Bayesian lognormal regression, we assessed the effects of evolocumab treatment on circulating metabolites such as lipoproteins and amino acids. 
In order to separate Lp(a) content from the LDL measurements, we applied a Gaussian mixture model to estimate the proportions of LDL particles with and without apolipoprotein(a) [apo(a)].

Results:
We identified that evolocumab treatment for 16 weeks resulted in 17% (95% credible interval: 8% to 26%) reduction of circulating Lp(a), coupled with substantial reduction of VLDL, IDL and LDL particles as well as their lipid contents.  
We also identified that about 24% [0.2% 47%] of large LDL particles were with apo(a). 
We did not observe changes in the proportion of LDL particles with apo(a) after treatment of evolocumab.

Conclusions:
Our analysis suggested that alternative pathways beyond LDL receptor mediated clearance may be involved in Lp(a) reduction by PCSK9 inhibition.

# Introduction

lipoprotein(a) [Lp(a)], synthesized in liver, is a low-density lipoprotein (LDL)-like particle with an apolipoprotein(a) [apo(a)] tail. 
Similar to other lipoproteins, Lp(a) is cleared from the circulation primarily by the liver [@Cain2005]. 
Epidemiological studies showed that subjects with elevated Lp(a) had increasing risk of developing cardiovascular disease [@ODonoghue2019]. 
Unlike LDL-cholesterol lowering, limited therapeutic options are available to reduce plasma Lp(a) concentration [@VanCapelleveen2016]. 
One of the agents that are capable to lower circulating Lp(a) levels is proprotein convertase subtilisin/kexin type 9 (PCSK9) monoclonal antibody, which prevents extracellular PCSK9 from binding to the LDL receptor, and results in supraphysiological abundance of LDL receptors [@Raal2014; @Raal2016].
The excess number of LDL receptors on the surface of hepatocytes were hypothesized to be responsible for the enhanced Lp(a) catabolism [@Romagnuolo2015; @Raal2016]. 
However, the above hypothesis does not concur with the effect of statins therapy, which significantly increases hepatic LDL receptor abundances but also increases plasma Lp(a) levels [@Tsimikas2019]. 
The same hypothesis was also challenged by the observation that LDL receptor knock out mice had similar Lp(a) catabolism as the wild type mice [@Cain2005]. 
Therefore, what roles LDL-receptor plays in Lp(a) catabolism has not been elucidated and remains a matter of debate [@Reyes-Soffer2017; @Watts2018; @Stoekenbroek2018]. 
In addition to the LDL receptor, very low density lipoprotein (VLDL) receptor and scavenger receptor class B type I (SR-BI) were also found to act as potential receptor for Lp(a), indicating more than one pathway is likely to be involved in Lp(a) reduction by PCKS9 inhibition [@Argraves1997; @Yang2013].
As a cost-effective approach, nuclear magnetic resonance (NMR) metabolomics simultaneously quantifies hundreds of circulating biomarkers and provides insights on multiple pathways [@Soininen2009].
In fact, NMR metabolomics has been used to study metabolic effects that were associated with  *LPA* loci, as well as genetic inhibition of PCSK9 [@Kettunen2016; @Sliz2018]. 
However, genetic inhibition of PCSK9 is different from inhibition with monoclonal antibody, and there is lack of studies on metabolic effects of PCSK9 inhibition with monoclonal antibody. 
In the current study, we aim to study metabolic changes corresponding to Lp(a) reduction in subjects with elevated Lp(a) after treatment of PCSK9 monoclonal antibody (evolocumab).
To this end, we performed NMR metabolomics of plasma samples derived from 30 healthy participants with elevated Lp(a) levels (placebo 14; evolocumab 16). 
Since Lp(a) and LDL significantly overlap in density, protein and lipid composition, LDL measurements made by current technologies contain the content of Lp(a) [@Yeang2015]. 
To address this issue, in this study we developed a statistical model to distinguish these two lipoproteins.

# Methods

## Study design

This was a site-specific sub study of the ANITSCHKOW trial (NCT02729025) conducted at the Amsterdam UMC in The Netherlands between April 2016 and July 2017. 
The ANITSCHKOW trial was a phase 3b, multicenter, randomized, double-blind, placebo-controlled trial using subcutaneous injections of evolocumab 420mg Q4W for 16 weeks as investigational product [@Stiekema2018].
The study was conducted in accordance to the principles of the Declaration of Helsinki and the study protocol was approved by the ethic committee of the Amsterdam UMC.

Patients were found to be eligible for enrollment in the ANITSCHKOW trial when they were $\geq$ 50 years of age, had a fasting LDL-C of $\geq$ 2.6 mmol/L (100mg/dL), and an Lp(a) of $\geq$ 125nmol/L (50mg/dL). 
At time of randomization, study subjects who were on lipid lowering treatment had to be on a stable dosing regimen for at least 8 weeks. 
Exclusion criteria were among others, diabetes mellitus and a cardiovascular event within three months prior to randomization. 
The complete list of eligibility criteria was reported previously [@Stiekema2018].

## Metabolite quantification

Blood samples were obtained at time of randomization and sixteen weeks after treatment. 
Patients were fasting for $\geq$ 9 hours for both blood withdrawals.

Lipoprotein(a) levels were measured using an isoform-independent immunoturbidometric assay (Polymedco, Cortlandt Manor, NY, USA) and reported in nmol/L.

Quantification of 225 metabolic measures was performed by using a high-throughput NMR metabolomics platform (Nightingale health, Finland) [@Soininen2009]. 
The 225 metabolic measures contain around 150 primary concentrations as well as ratios that were derived from the primary concentrations. 
In this study, we focused on primary concentrations of circulating metabolites that cover multiple metabolic pathways including lipoproteins, fatty acids as well as amino acids and glycolysis intermediates. 
The following 14 lipoprotein subclasses and their lipid compositions were quantified: 
extremely large (average particle diameter >75 nm), 
very large (average particle diameter 64.0 nm), large (53.6 nm), medium (44.5 nm), small (36.8 nm) and very small VLDL (31.3 nm); 
intermediate density lipoprotein (IDL; 28.6 nm); 
three LDL subclasses, i.e. large (25.5 nm), medium (23.0 nm) and small LDL (18.7 nm); 
and four HDL subclasses, i.e. very large (14.3 nm), large (12.1 nm), medium (10.9 nm) and small HDL (8.7 nm).
The complete list of the 225 metabolic measures can be found at  https://nightingalehealth.com/biomarkers

## Statistical analysis

### Metabolic effects of PCSK9 inhibition with evolocumab

To assess the effect of evolocumab on levels of a circulating metabolite, we ran lognormal regression analysis. 
The outcome variable ($y$) was concentration of a metabolite. 
Two predictor variables were in the regression model: 1) evolocumab treatment (evolocumab, $T_i$ = 1; placebo $T_i = 0$); 
2) metabolite concentration at time of randomization ($x$).
The baseline concentrations were centered and scaled so that the mean is 0 and standard deviation is 1.
The lognormal distribution was chosen to model the outcome variable because its values were positive continuous. 
Due to missing observations in outcome variables, we applied the Bayesian approach to handle the missing data. 
There are two types of missing values, which are as follows: 
(1) when the concentration of a metabolite is below the limit of detection, 
or (2) when values were rejected by the automatic sample and measurement quality control procedure in the Nightingale pipeline. 
All the missing observations were assumed missing at random and treated as parameters.
Values were randomly drawn from a lognormal distribution. 
Regarding the missing values that were below the limit of detection, the imputed values were constrained between zero and the minimal observed value. 
We fitted the model by running Hamiltonian Markov Chain Monte Carlo in the program Stan (version 2.18.3).
We ran four Markov chains with 2000 iterations in each chain. 
Results were presented with the posterior mean of lognormal regression coefficient with 95% credible interval (CI). 
The lognormal regression coefficient can be interpreted as the expected proportional change in the metabolite concentration.

### Separating Lp(a) content from LDL measures

Since Lp(a) and LDL significantly overlap in density, as well as protein and lipid composition, LDL measures made by current technologies contain the content of Lp(a). 
In this study, we applied the Gaussian mixture model to address this issue. 
Our statistical model was inspired by the biological relationship between Lp(a) and LDL. 
Lp(a) is often considered to comprise a LDL particle and covalently bound apo(a). 
The diameter of Lp(a) is about 25 nm [@BERG1963]. 
Based on the above knowledge, we assumed that the pool of large LDL particle (average diameter 25.5 nm) is a mixture of LDL particles with and without apo(a). 
The outcome variable ($y$) was the particle concentration of large LDL, and described by a mixture of two normal distributions with a proportion parameter $\lambda$.

$$P(y|\lambda, \mu, \sigma) = \prod_{i = 1}^{N}(\lambda \times N(\mu_1,\sigma_1) + (1 - \lambda) \times N(\mu_2, \sigma_2))$$

The parameter $\mu_1$ and $\sigma_1$ were estimated based on the particle concentration of Lp(a), which was measured by using the isoform-independent immunoturbidometric assay. 
The parameter $\mu_2$ was modeled as $\mu_2[i] = \alpha_L + \beta_{S,L} \times S[i] + \beta_{E,L} \times E[i] + (\beta_L + \beta_{P,L} \times P[i]) \times V[i]$.

There were four covariates $V$, $S$, $E$ and $P$. 
$V = 0$ referred the measurement was made at time of randomization, whereas $V = 1$ refereed to sixteen weeks after treatment. 
Because lipid-lowering medication such as statins, ezetimibe as well as evolocumab has causal effects on these LDL particle concentrations, we included three covariates (statin usage: $S = 1$; Ezetimibe usage $E = 1$; Evolocumab usage $P = 1$). 
Regarding parameter $\lambda$,  proportion of LDL particles with apo(a) tail, we set the beta distribution $Beta(1,1)$ as its prior distributions.

# Results

The current study included 30 patients (evolocumab, n = 16; placebo, n = 14). 
Baseline characteristics of the two groups were comparable and provided in Table 1.

## Metabolic effects of PCSK9 inhibition with evolocumab

Evolocumab treatment for 16 weeks resulted in mean 17% (95% credible interval: 8% to 26%) reduction in Lp(a) compared with placebo.

To identify the metabolic effects corresponding to the Lp(a) reduction by PCSK9 inhibition, we performed NMR metabolomics covering metabolic pathways such as lipoproteins, fatty acids, amino acids and glycolysis. 
We observed that evolocumab treatment resulted in substantial reduction in particle concentration of extremely large (81% [51% 100%]), very large (89% [69% 100%]), large (61% [36%, 86%]), medium (50% [36%, 63%]), small (39% [31% 46%]) and very small VLDL (47% [41% 54%]) compared with the placebo group.
We also observed that evolocumab treatment resulted in particle concentration reduction in IDL (53% [45%, 61%]), large (57% [49% 66%]), medium (59% [50%, 6%]) and small LDL (55% [47%, 64%]) compared with the placebo group. 
In addition, we observed that evolocumab treatment resulted in decreased concentrations of very large HDL particles (25% [4% 46%]), and increased concentrations of medium HDL particles (13% [4% 23%]) (Figure 1).

![Figure 1 Mean difference between evolocumab and palacebo group, adjusting for pre-treatment lipoprotein particle concentrations. Circles represent the posterior means. Lines refer to the 95% credible intervals](./Treatment_effect_lipoproteins.png)

Similar to the lipoprotein particle concentration profiles, we observed that evolocumab treatment resulted in substantial reduction of esterified cholesterol (CE), triglyceride (TG), free cholesterol (FC) and phospholipid (PL) in VLDL, IDL and LDL as well as very large HDL, compared with the placebo group (Figure 2).

![Figure 2 Mean difference in lipoprotein lipid compositions between evolocumab and placebo group, adjusting for pre-treatment lipoprotein lipid concentrations. Circles represent the posterior means. Lines refer to the 95% credible intervals. CE: esterified cholesterol; FC: free cholesterol; PL: phospholipids; TG: triglyceride.](./Treatment_effect_lipoprotein_lipid.png)

The NMR metabolomics also quantified the fatty acid content in the lipoproteins. 
We observed that treatment of evolocumab resulted in 30% [24% 36%] reduction in total fatty acids, with the largest effect on docosahexaenoic acid (DHA 40% [24% 55%]) (Figure 3).

![Figure 3 Mean difference in fatty acid concentrations between evolocumab and placebo group, adjusting for pre-treatment fatty acid concentrations. Circles represent the posterior means. Lines refer to the 95% credible intervals.](./Treatment_effect_FFAs.png)

We observed no difference in concentrations of other metabolites covering amino acids, fluid balance, glycolysis and ketone bodies.

## Separating Lp(a) content from the LDL measures

Lp(a) is thought to be comprised of a LDL particle and a covalently bound apo(a). 
Because Lp(a) and LDL significantly overlap in density, as well as protein and lipid composition, LDL measures contain the content of Lp(a). 
In this study, we focused on the large LDL particle, because its diameter (25.5 nm) is close to the diameter of Lp(a). 
At baseline, we identified that on average 24% ([0.2%, 47%]) large LDL particles were with apo(a) (Figure 4).

Sixteen weeks after treatment of evolocumab, on average 12% (0.02%, 36%) large LDL particles were with apo(a) in the placebo group. 
In the treatment group, the average proportion of large LDL particles with apo(a) was 9% (0.0008%, 31%).
We did not observe that the proportions of large LDL particles with apo(a) were different between the treatment and placebo group (-3% [-36% 23%]).

# Discussion

The major finding of this work is that evolocumab treatment in patients with high Lp(a) massively decreases large VLDL in addition to the well known effect on LDL. 
In the current work, evolocumab treatment resulted in on average 17% reduction in plasma Lp(a) concentration compared with placebo. 
Our detailed metabolomic profiling revealed that Lp(a) reduction by PCSK9 inhibition corresponded to substantial reduction in concentrations of VLDL, IDL and LDL particles as well as their lipid and fatty acid contents. 
Our observation is consistent with the previous study, in which evolocumab treatment not only reduced concentrations of LDL particles, but also substantially reduced triglyceride-rich lipoprotein particles [@Toth2018]. 
Since a PCSK9 monoclonal antibody such as evolocumab reduces the degradation of the LDL receptor as well as the VLDL receptor, the very low concentration of VLDL particles after the treatment may be ascribed to the increased number of VLDL receptors [@Poirier2008]. 
The upregulation of VLDL receptors can enhance uptake of VLDLs but also Lp(a) from the circulation [@Argraves1997]. 
In addition, our analysis also showed that treatment of evolocumab lowered concentrations of very large HDL particles. 
Very large HDL is cholesterol enriched and mainly captured by the SR-B1 present in the liver [@Acton1996]. 
Similar to the VLDL receptor, SR-B1 has also been shown to bind Lp(a), and its upregulation will probably enhance the Lp(a) catabolism [@Yang2013].

In our analysis, we observed a large discordance in Lp(a) and LDL reduction in response to evolocumab. 
This discordance was suggested by a recent study to be related to alternative pathways of Lp(a) reduction beyond LDL receptor mediated clearance [@Shapiro2019]. 
Lp(a) and LDL significantly overlap in density, as well as protein and lipid composition. 
As a result, current technologies including the NMR metabolomics used in this study cannot distinguish these two lipoproteins, and include the content of Lp(a) in the LDL measurements. 
We think that it is important to separate the Lp(a) content from the LDL measurements in order to better understand the discordance between these two lipoproteins in response to evolocumab. 
To this end, we applied a Gaussian mixture model based on the assumption that Lp(a) was comprised of a LDL particle and a covalently bound apo(a). 
We identified that on average about 24% of large LDL particle had apo(a) tail. 
We did not observe changes in proportions of large LDL particle with apo(a) tail due to evolocumab treatment.

## Strengths and limitations

Strengths and limitations of our study warrant consideration. 
Our study provided the first NMR metabolomics data for a PCSK9 inhibition trial. 
The detailed metabolomic profiling revealed metabolic effects of the PCSK9 inhibitor, and improved our understanding of this lipid-lowering therapy. 
Moreover, in this study we developed a statistical method to separate Lp(a) content from the LDL measurements. 
However, due to the heterogeneity of Lp(a) and limited number of participants, our estimation of the proportions of LDL with apo(a) had very large uncertainty. 
Applying our statistical model with a larger number of participants as well as validation in the laboratory are warranted in the future studies.

In conclusion, our NMR metabolomic profiling revealed that Lp(a) reduction by PCSK9 inhibition corresponded to substantial reduction in VLDL, IDL and LDL particles as well as their lipid contents. 
Our analysis suggested that alternative pathways beyond LDL receptor mediated clearance may be involved in Lp(a) reduction by PCSK9 inhibition.

# Authors

# Acknowledgements

# Sources of Funding

# Disclosures

# Affiliations

# References